---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

```{r}
library(dplyr)
library(ggplot2)
library(scales)
```


## What is mean total number of steps taken per day?

```{r}
actDat <- read.csv("./activity.csv", colClasses = c("numeric", "Date", "numeric"))
actDay <- actDat %>% group_by(date) %>% summarize(Steps = sum(steps,na.exclude = TRUE))
p1 <- ggplot(data = actDay, mapping = aes(x = Steps)) + geom_histogram(fill = "red", 
    colour = "black") + scale_x_continuous("Steps per Day", labels = comma) + 
    scale_y_continuous("Frequency") + ggtitle("Total Number of Steps Taken Each Day")
p1
```
## The mean and median of the total number of steps taken per day are:
```{r}
mean(actDay$Steps,na.rm = TRUE)
median(actDay$Steps,na.rm = TRUE)
```

## What is the average daily activity pattern?
```{r}

step_average <- actDat %>% group_by(interval) %>% summarize(average_step = mean(steps,na.rm = TRUE))

p2 <- ggplot(data = step_average,mapping = aes(x = interval, y = average_step))+geom_line()+ scale_x_continuous("Day Interval", breaks = seq(min(step_average$interval), 
    max(step_average$interval), 100)) + scale_y_continuous("Average Number of Steps") + 
    ggtitle("Average Number of Steps Taken by Interval")
p2
```

## Which day has the maximium average number of steps 
```{r}
max_interval <- which.max(step_average$average_step)
actDat[max_interval,]
```
We have 288(60*24/5) intervals per day in total, so the max happens at 104th interval, around 8:35AM. 

## Imputing missing values

### Total number of missing values in the dataset are: 
```{r}
sum(is.na(actDat$steps) )
```

### Use mean/median for that day to replace all NAs in that day
```{r}
actDat2 <- actDat %>% left_join(step_average, by = "interval")
actDat2$steps <- ifelse(is.na(actDat2$steps), actDat2$average_step, actDat2$steps)
actDat2 <- select(actDat2,-average_step)

```

### repeate analysis above. It appears that the frequency plot shows some differences although mean/median does not change significantly
```{r}
actDay <- actDat2 %>% group_by(date) %>% summarize(Steps = sum(steps))
p3 <- ggplot(data = actDay, mapping = aes(x = Steps)) + geom_histogram(fill = "red", 
    colour = "black") + scale_x_continuous("Steps per Day", labels = comma) + 
    scale_y_continuous("Frequency") + ggtitle("Total Number of Steps Taken Each Day")
p3

mean(actDay$Steps,na.rm = TRUE)
median(actDay$Steps,na.rm = TRUE)
```

## Are there differences in activity patterns between weekdays and weekends?
```{r}
library(lubridate)
actDat2$weekend <- ifelse(wday(actDat2$date) %in% c(6:7),"weekend","weekday")

actDay <- actDat2 %>% group_by(interval,weekend) %>% summarize(average_step = mean(steps,na.rm = TRUE))
p4 <- ggplot(data = actDay,mapping = aes(x = interval, y = average_step))+geom_line()+facet_grid(.~weekend)+ scale_x_continuous("Day Interval", breaks = seq(min(actDay$interval), 
    max(actDay$interval), 100)) + scale_y_continuous("Average Number of Steps") + 
    ggtitle("Average Number of Steps Taken by Interval")
p4
```

